---
title: "08 Machine Learning - Course Project"
author: "Technophobe01"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document:
    fig_caption: no
    number_sections: yes
header-includes: \usepackage[compact]{titlesec} \titlespacing{\section}{0pt}{0pt}{0pt}
---

```{r setOptions, echo=FALSE, message=FALSE, results='hide'}
require(knitr)
opts_chunk$set(fig.width=8, fig.height=5, 
               fig.path='Figs/',
# Dev Setup / Production setup
#               cache=FALSE, echo=TRUE, warning=TRUE, message=TRUE, results='show')
               cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, results='show')
require(knitcitations)
cleanbib()
```

```{r Setup, results='hide'}
setwd("~/Documents/Dropbox/dev/cousera/courses/08_PracticalMachineLearning/00_Course_Work/CourseProject")

# Load the required R packages...
#
requiredPackages <- c("ggplot2",        # Used to plot graphics 
                      "gtable",
                      "ggthemes",       # Extra themes, scales and geoms for ggplot (Very cool!)
                      "xtable",
                      "caret",          # Classification and Regression Training Library
                      "randomForest")   # Breiman and Cutler's random forests for classification and regression

ipak <- function(pkg){
        new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
        if (length(new.pkg))
                install.packages(new.pkg, dependencies = TRUE)
        sapply(pkg, require, character.only = TRUE)
}

ipak(requiredPackages)
```

# Executive Summary  

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the _**quantified self movement**_ – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. 

One thing that people regularly do is quantify how much of a particular activity they do, _but_ **they rarely quantify how well they do it**. 

## Report Goal

The goal of this project is to use the referenced data caputured from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to *predict* how the participants performed the exercises. This report describes: 

1. How a model was built to frame and analyse the data
1. How cross validation was used to confirm the results and what the estimated expected out of sample error is of the analysis
1. What choices were made to define the results and guide the anlysis.

seperaralty we use the prediction data to generate 

1. 20 different test case results which are to be submitted to coursera. 

**Note:** More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset) on the specific exercises measured and predicted. 

## Data Set Description

Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg).

The participants were asked to perform barbell lifts *correctly* and *incorrectly* in 5 different ways. Six young healthy participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: 

1. (Class A) - exactly according to the specification 
1. (Class B) - throwing the elbows to the front 
1. (Class C) - lifting the dumbbell only halfway 
1. (Class D) - lowering the dumbbell only halfway 
1. (Class E) - throwing the hips to the front 


Read more: http://groupware.les.inf.puc-rio.br/har#dataset#ixzz3gge1uKYv

## Data Processing

``` {r loadData}

###############################################################################
#
#  1. Load, the training and test sets 
#
# The first step is to download the training and test data into the data
# subdirectory, that we create if it does not exist. This program assumes that
# all the data will be dowloaded into the 'data' directory below the working
# directory.
#

trainDataURL = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testDataURL = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

dataDir         <- "./data"
trainData       <- "pml-training.csv"
testData        <- "pml-test.csv"

trainfilePath   <- file.path(dataDir, trainData)
testDataPath    <- file.path(dataDir, testData)

# If the data dire does not exist we create it, so it can be used to store the
# downloaded data

if (!file.exists(dataDir)) {
        dir.create(dataDir)
}

# Now we can download files into the data subdirectory. 

if (!file.exists(trainfilePath)) {
        download.file(trainDataURL, temp, mode="wb", method = "curl")
        unlink(temp)
}

if (!file.exists(testfilePath)) {
        download.file(testDataURL, temp, mode="wb", method = "curl")
        unlink(temp)
}

```

# Analysis Goals

## Analysis Results

# Conclusions

\pagebreak

# Appendix A


\pagebreak


# **References**

- *Data Classification of Body Postures and Movements.* 

    - By: Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' 
    - Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 

- *Applied Predictive Modeling*   

    - By: Max Kuln, Kjell Johnson, Publisher: Springer; Pub Date: May 17, 2013, eISBN: 978-1-461-46848-6

- *The R Book*   

    - By: Michael J. Crawley, Publisher: John Wiley & Sons Pub. Date: December 26, 2012, eISBN: 978-1-118-44896-0   
    
- *An Introduction to Statistical Learning*

    - By: Gareth James (Author), Daniela Witten (Author), Trevor Hastie (Author), Robert Tibshirani (Author)

- *The Elements of Statistical Learning*

    - By: Trevor Hastie (Author), Robert Tibshirani (Author), Jerome Friedman (Author)

- *R in Action*    

    - By: Robert Kabacoff Publisher: Manning Publications Pub. Date: August 24, 2011, ISBN-10: 1-935182-39-0   

- *Mathematical Statistics with Resampling and R*

    - By: Laura Chihara; Tim Hesterberg Publisher: John Wiley & Sons Pub. Date: September 6, 2011 Print ISBN: 978-1-11-02985-5   

- *OpenIntro Statistics - 2nd Editions*

    - By: by David M Diez (Author), Christopher D Barr (Author), Mine Çetinkaya-Rundel (Author) 

- *Think Stats, 2nd Edition*

    - By: Allen B. Downey Publisher: O'Reilly Media, Inc. Pub. Date: October 28, 2014 Print ISBN-13: 978-1-4919-0733-7

[0]: https://class.coursera.org/predmachlearn-030
[1]: http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html
[2]: https://github.com/Technophobe01/courses/tree/master/08_RegressionModels/00_Course_Work/CourseProject

