---
title: "08 Machine Learning - Course Project"
author: "Technophobe01"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document:
    fig_caption: no
    number_sections: yes
header-includes: \usepackage[compact]{titlesec} \titlespacing{\section}{0pt}{0pt}{0pt}
---

```{r setOptions, echo=FALSE, message=FALSE, results='hide'}
require(knitr)
opts_chunk$set(fig.width=8, fig.height=5, 
               fig.path='Figs/',
# Dev Setup / Production setup
#               cache=FALSE, echo=TRUE, warning=TRUE, message=TRUE, results='show')
               cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, results='show')
require(knitcitations)
cleanbib()
```

```{r Setup, results='hide'}
setwd("~/Documents/Dropbox/dev/cousera/courses/08_PracticalMachineLearning/00_Course_Work/CourseProject")

# Load the required R packages...
#
requiredPackages <- c("ggplot2",        # Used to plot graphics 
                      "gtable",
                      "ggthemes",       # Extra themes, scales and geoms for ggplot (Very cool!)
                      "xtable",
                      "caret",          # Classification and Regression Training Library
                      "randomForest")   # Breiman and Cutler's random forests for classification and regression

ipak <- function(pkg){
        new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
        if (length(new.pkg))
                install.packages(new.pkg, dependencies = TRUE)
        sapply(pkg, require, character.only = TRUE)
}

ipak(requiredPackages)
```

# Executive Summary  

## Data Set Description

## Data Processing

# Analysis Goals

## Analysis Results

# Conclusions

\pagebreak

# Appendix A


\pagebreak


# **References**

- *Applied Predictive Modeling*   

    - By: Max Kuln, Kjell Johnson, Publisher: Springer; Pub Date: May 17, 2013, eISBN: 978-1-461-46848-6

- *The R Book*   

    - By: Michael J. Crawley, Publisher: John Wiley & Sons Pub. Date: December 26, 2012, eISBN: 978-1-118-44896-0   
    
- *An Introduction to Statistical Learning*

    - By: Gareth James (Author), Daniela Witten (Author), Trevor Hastie (Author), Robert Tibshirani (Author)

- *The Elements of Statistical Learning*

    - By: Trevor Hastie (Author), Robert Tibshirani (Author), Jerome Friedman (Author)

- *R in Action*    

    - By: Robert Kabacoff Publisher: Manning Publications Pub. Date: August 24, 2011, ISBN-10: 1-935182-39-0   

- *Mathematical Statistics with Resampling and R*

    - By: Laura Chihara; Tim Hesterberg Publisher: John Wiley & Sons Pub. Date: September 6, 2011 Print ISBN: 978-1-11-02985-5   

- *OpenIntro Statistics - 2nd Editions*

    - By: by David M Diez (Author), Christopher D Barr (Author), Mine Çetinkaya-Rundel (Author) 

- *Think Stats, 2nd Edition*

    - By: Allen B. Downey Publisher: O'Reilly Media, Inc. Pub. Date: October 28, 2014 Print ISBN-13: 978-1-4919-0733-7

[0]: https://class.coursera.org/predmachlearn-030
[1]: http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html
[2]: https://github.com/Technophobe01/courses/tree/master/08_RegressionModels/00_Course_Work/CourseProject

